A scoping review of rodent trapping studies to understand the prevalence of zoonotic disease in West Africa.

### 2021-10-05 reports/manuscript_v2.Rmd
need to sort figures

---
title: "A scoping review of rodent trapping studies to understand the prevalence of zoonotic disease in West Africa."
description: |
date: "`r Sys.Date()`"
author: 
  - first_name: "David"
    last_name: "Simons"
    url: https://www.dsimons.org
    orcid_id: 0000-0001-9655-1656
output: 
  # distill::distill_article:
  #   toc: true
  #   toc_depth: 3
  #   toc_float: true
  # word_document:
  pdf_document:
    toc: true
    toc_depth: 3
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here())
  })
always_allow_html: yes
bibliography: "`r here::here('citations', 'manuscript_citations.bib')`"
---
```{r, echo=FALSE, include=FALSE}
source(here::here("scripts", "manuscript_library.R"))
```

@@

The potential effect of zoonotic infectious diseases outbreaks on human health and societies has been dramatically highlighted through the ongoing SARS-CoV-2 pandemic and recent Ebola virus outbreaks. The number of zoonotic spillover events are projected to increase under intensifying anthropogenic pressure such as, increased human populations [@allen_global_2017], increasing urbanisation [@hassell_urbanization_2017], and global climate change [@morse_prediction_2012]. In addition, two taxa are proposed to contribute to the greatest number of potential zoonotuc pathogens globally, rodents (Rodentia) and bats (Chiroptera) [@han_rodent_2015]. Of 2,220 extant rodent species, 244 (10.7%) are described as being reservoirs of 85 zoonoses [@han_global_2016]. Specifically, West Africa has previously been identified as region at increased hazard for rodent-borne zoonotic spillover events under different projected scenarios [@garcia-pena_land-use_2021]. Rodents are implicated in the transmission pathways of several human diseases in this region, including, Lassa fever, Schistosomiasis and Leptospirosis.

Rodent species form diverse assemblages, which provide important and beneficial ecosystem services including pest regulation and seed dispersal [@fischer_ecosystem_2018]. The role of rodent species' in zoonotic infectious disease spillover or novel pathogen emergence can be viewed as an ecosystem disservice. Rodents typically demonstrate “fast” life histories [@dobson_fast_2007], these traits are also associated with a species being a reservoir of zoonotic pathogens [@han_rodent_2015; @albery_fast-lived_2021]. Further, these traits are prevalent in species that thrive in human dominated landscapes where they displace species that are less likely to be reservoirs of zoonotic pathogen [@gibb_zoonotic_2020]. The widespread occurrence of reservoir species and their proximity to human activity make the description of rodent species assemblages and host-pathogen interactions vitally important to understanding the risk of zoonotic spillover and novel pathogen emergence.

@@

These studies have been used to identify potential geographic hotspots where virus and host species diversity may be expected to be at its greatest to predict regional zoonotic disease spillover risks. However, there remains the potential for important confounding in these spatial distributions through bias generated by study design and selection of sampling sites. For example, systematically increased sampling (e.g. more intensive studies over longer time periods) or over-representation of certain habitats (e.g. peri-urban landscapes) could lead to an apparent association between locations and risk that is driven by these factors rather than an underlying host and virus association [@wille_how_2021; @gibb_mammal_2021]. Conversely some regions may not be sampled adequately and therefore under-represented in these datasets due to sparse human populations or inaccessible habitats leading to these regions being reported to be at low risk of zoonotic disease spillover events.

@@

We extracted data from eligible studies into a Google sheets document. Extracted variables included i) study identifiers; ii) study aims; iii) trapping methodology; iv) geolocation data; v) method of speciation; vi) trapping locations and dates; vii) trapped species; viii) number of trap-nights and ix) microorganisms/pathogens of interest. The data extraction tool is archived and available in Supplementary Material 1.

@@

We summarised the number of studies, the year in which trapping occurred and the country in which they were conducted. We used the GPS coordinates of single trap, trapping grid/line or study site and the number of trap nights to calculate trapping effort (trap night density) within level 2 administrative areas in West Africa. For studies not reporting the number of trap nights we imputed the number based on the median trap success rate from studies which reported the number of trap nights (matched to building or non-building based study sites). The median trap success rate for all rodents at a defined trap site was calculated separately for trap sites which included built environments and non-built environments. The number of rodents trapped at a trap site was then used in combination with these two values of trap success to impute the number of trap nights for trap sites with no reported trap nights. We summarised the habitat types of trap sites based on information reported in the study.

For studies investigating rodent zoonoses we compared the location of trapping sites with SEDAC Global Population Density estimates for 2005 - the median year studies were commenced. We used a Generalised Additive Model (GAM) incorporating a spatial interaction term to investigate the association of number of trap nights and human population density. The model structure was specified as:

  + *Trap night density* ~ *Tweedie*(log(Population density (2005)) + (Longitude * Latitude))

We obtained land cover classifications from the European Space Agency Copernicus dataset at 300m^2^ spatial resolution (2005) we extracted the proportion of land cover classes within all regions of West Africa and all regions in which rodent trapping occurred for investigation of zoonotic diseases. We compared the proportion of possible land cover classes with those of regions where trapping occurred.

Finally, we mapped the presence and absence of rodent species and compared this to the presence and absence reported by both GBIF and IUCN to give a measure of the extent of each species range in which they have been sampled.

@@

We identified 4,282 relevant citations, with 126 rodent trapping studies included. The earliest included studies were published in 1974 with increasing numbers of studies being performed annually since 2005. The median length of rodent trapping activity was one year (IQR 0-2 years). The median time from completion of rodent trapping to publication is 3 years (IQR 2-5 years) (Figure 1.).

```{r, figure_1, echo=FALSE, out.width="100%"}

if(knitr::is_html_output()) {
  knitr::include_graphics(here("figures", "Fig_1_Panel_A.png"), dpi = 320)
} else {
  knitr::include_graphics(here("figures", "Fig_1_Panel_A.png"))
}

*Figure 1*: Each row represents one of the 126 included studies, green points designate the first year of data collection, blue points designate the end of data collection. For studies completed within one year the blue point completely overlies the green. Studies with a transparent grey point did not report the year in which trapping was conducted. The year of publication is shown by a red point.

@@

@@

```{r, figure_2, layout = "l-screen-inset", echo=FALSE, out.width="100%"}

knitr::include_graphics(here("figures", "Figure_2.png"))

*Figure 2*: **Panel A:** Map of West Africa, countries where rodent trapping has occurred are mapped to level 2 administrative areas (where available). These regions are coloured by the total number of trap nights performed at trap sites within their boundaries. **Panel B:** (n.b. will not include map in final manuscript) The number of trap nights conducted is associated with a regions population density. The population density for all regions in West Africa is shown in the yellow box plot, the population density and trap night density for each region is show on the purple scatterplot. The line of best fit is a GAM model not incorporating spatial interactions. The map panel on the right is the product of the GAM model incorporating spatial interactions. **Panel C:** For each of the 10 land cover classes from the ESA dataset we measure the proportion of 300m^2^ pixels within a level 2 administrative area. Zoonotic trapping studies occurred in regions over-representative for Cropland, Mosaic landscapes and Shrubland while being under-representative for Bare and Sparse vegetation land cover classes.

@@

```{r, figure_3, layout = "l-screen-inset", echo=FALSE, out.width="100%"}

knitr::include_graphics(here("figures", "Figure_3.png"))

*Figure 3*: Each row corresponds to a single rodent species. The column on the left shows the presence and absence of a rodent species from the individual studies included in this review. The centre column shows the presence of a rodent species obtained from GBIF (September 2021) for records where longitude and latitude have been provided. The right column shows the range of rodent species as proposed by the IUCN (2021) (red shaded area), overlaid are the presence points from both this review and GBIF records.

@@

```{r, figure_4, echo=FALSE, out.width="100%"}

knitr::include_graphics(here("figures", "Figure_4a.png"))
knitr::include_graphics(here("figures", "Figure_4b.png"))
knitr::include_graphics(here("figures", "Figure_4_legend.png"))

@@

*Figure 4*: Presence/absence plots at each unique trapping site for the four most commonly assayed microorganisms Arenaviridae (top), *Bartonella sp.* (second row), *Borrelia sp.* (third row) and *Toxoplasma gondii* (fourth row). The tables to the right of each map highlight the 10 (or number applicable) most commonly positive and tested rodent species and genera assayed.

@@

```{r figure_5, echo=FALSE, out.width="100%"}

knitr::include_graphics(here("figures", "Figure_5.png"))

*Figure 5*: The percentage of assays from each included species and microorganism tested for is represented by colour. The size of a point relates to the squareroot of the number of individuals assayed. The plot on the left is for species/genera assayed for four or more microorganisms, the plot on rights are for species/genera assayed for 3 or less microorganisms.

# Discussion    
We identified 124 rodent trapping studies conducted in the West African region since 1965. These studies were conducted in 14 countries with 54 studies providing information about zoonotic pathogens in trapped rodents. Studies were conducted in all West African countries except Gambia and Togo with between- and within-country variability in the number of trap nights conducted. Over 70,000 rodents identified to genus or species level were trapped from 112 species of Rodentia. Thirty-two potential pathogens were investigated across all studies, 5 rodent species were assayed for at least 18 potential pathogens. We identify several important limitations to available data, potentially biasing previous zoonotic risk studies, first, in the timing of studies, and the locations in which they've been conducted, second, the locations of species presence and third, the selection of pathogens to assay.

First, timing and location of trapping studies can introduce bias to inference from these data when investigating the risk of zoonotic disease emergence. Most studies have been conducted within the last twenty years. Human population, intensification of agriculture and urbanisation have proceeded rapidly over this time across West Africa. These changes modify the available habitats for rodent species increasing the prevalence of commensal species that are more likely to be reservoirs of zoonoses. Not accounting for the time period in which data were collected will introduce bias into inferences made for zoonotic emergence based on reservoir species distributions. Further, several regions, namely North West Senegal, Eastern Sierra Leone, Southern Ghana, Southern Benin, South West Nigeria  Central Nigeria have been sampled at much higher intensity than elsewhere in West Africa. Studies modelling risk of zoonoses without incorporating this important sampling heterogeneity will introduce bias when inferring current or projected risk.

Second, rodent species were found to have wider distributions than the ranges reported by IUCN data. Trapping studies were conducted from more regions across West Africa than were available from GBIF records and where therefore able to report presence and absence of species at greater spatial resolution. Further, the additional data produced from species' absence and pseudoabsence highlighted the variability of rodent presence within their ranges across the region. Failing to incorporate this heterogeneity in rodent reservoir species' distribution can importantly bias estimations of the risk of zoonotic disease emergence. For example, the apparent absence of *Mastomys natalensis* at several trapping sites in Nigeria may explain some of the observed spatial variation in Lassa fever spillover that would not be expected based on it's IUCN proposed range or presence only data from GBIF.

Third, while up to 32 potential zoonoses were investigated only four were investigated in more than three studies. Of these four studies conducting assays for *Lassa mammarenavirus* or other Arenaviridae where the most widely conducted in regards to number of countries and number of study sites within countries. For example, *Toxoplasma sp.* were found to be prevalent in rodent hosts within Senegal and despite a significant burden of human disease no studies investigated the prevalence of this pathogen in other countries of West Africa [https://www.nature.com/articles/s41598-020-69078-9]. The limited geographic scope of sampling for these pathogens hampers efforts to understand the risk of zoonotic spillover across this heterogeneous region.

Finally, we found that only five of the 112 rodent species trapped had been tested for more than half of the potential pathogens. There is increasing emphasis on the role of co-infection and multi-species pathogen transmission systems on zoonotic disease spillover risk which depends on comprehensive testing of both known and un-known reservoir species. More comprehensive testing of potential reservoir species across West Africa is required to better understand the role of individual species in disease systems. This will also reduce the risk of bias in predictive risk modelling by better characterising host-pathogen associations in species rich environments.

@@

We have identified important biases in the locations of studies, the trapping effort for different rodent species and the selection of zoonotic pathogens to assay across West Africa. The use of these data to infer risk of zoonotic spillover or novel disease emergence is likely inappropriate unless these systematic biases are addressed. We hope that the current synthesis of extant data will encourage further comprehensive ecological studies on rodent species assemblages and their pathogens. One priority would be to better understand the true distribution of known rodent reservoirs (particularly *Mastomys natalensis*) across the region with particular emphasis made to the habitat in which they reside and the structure of rodent assemblages in the region.

@@

>>>>

---
title: "A scoping review of rodent trapping studies to understand the prevalence of zoonotic disease in West Africa."
date: "`r Sys.Date()`"
author: 
  - "David Simons"
    # url: https://www.dsimons.org
    # orcid_id: 0000-0001-9655-1656
output: 
  # distill::distill_article:
  #   toc: true
  #   toc_depth: 3
  #   toc_float: true
  # word_document:
  pdf_document:
    toc: true
    toc_depth: 3
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here())
  })
always_allow_html: yes
bibliography: "`r here::here('citations', 'manuscript_citations.bib')`"
---
```{r, echo=FALSE, include=FALSE}
source(here::here("scripts", "manuscript_library.R"))
```

@@

The potential effect of zoonotic infectious diseases outbreaks on human health and societies has been dramatically highlighted through the ongoing SARS-CoV-2 pandemic and recent Ebola virus outbreaks. The number of zoonotic spillover events are projected to increase under intensifying anthropogenic pressure such as, increased human populations [@allen_global_2017], increasing urbanisation [@hassell_urbanization_2017], and global climate change [@morse_prediction_2012]. In addition, two taxa - rodents (Rodentia) and bats (Chiroptera) - are proposed to contribute to the greatest number of zoonotic pathogens and the greatest risk of novel pathogen emergence [@han_rodent_2015]. Of 2,220 extant rodent species, 244 (10.7%) are described as being reservoirs of 85 zoonoses [@han_global_2016]. Specifically, West Africa has previously been identified as region at increased hazard for rodent-borne zoonotic spillover events under different projected scenarios [@garcia-pena_land-use_2021]. Rodents are implicated in the transmission pathways of several human diseases in this region, including, Lassa fever, Schistosomiasis and Leptospirosis [@meerburg_rodent-borne_2009].

Rodent species form diverse assemblages, which provide important and beneficial ecosystem services including pest regulation and seed dispersal [@fischer_ecosystem_2018]. The role of rodent species' in zoonotic infectious disease spillover or novel pathogen emergence can be viewed as an ecosystem disservice. Rodents typically demonstrate “fast” life histories [@dobson_fast_2007], these traits are also associated with a species being a reservoir of zoonotic pathogens [@han_rodent_2015; @albery_fast-lived_2021]. Further, these traits are prevalent in species that thrive in human dominated landscapes where they displace species that are less likely to be reservoirs of zoonotic pathogen [@gibb_zoonotic_2020]. The widespread occurrence of reservoir species and their proximity to human activity make the description of rodent species assemblages and host-pathogen interactions vitally important to understanding the risk of zoonotic spillover and novel pathogen emergence [@han_global_2016].

@@

These studies have been used to identify potential geographic hotspots where virus and host species diversity may be expected to be at its greatest to predict regional zoonotic disease spillover risks. However, there remains the potential for important confounding in these spatial distributions through bias generated by study design and selection of sampling sites [@plowright_sampling_2019]. For example, systematically increased sampling (e.g. more intensive studies over longer time periods) or over-representation of certain habitats (e.g. peri-urban landscapes) could lead to an apparent association between locations and risk that is driven by these factors rather than an underlying host and virus association [@wille_how_2021; @gibb_mammal_2021]. Conversely some regions may not be sampled adequately and therefore under-represented in these datasets due to sparse human populations or inaccessible habitats this may result in these regions being reported inappropriately as at low risk of novel pathogen emergence or zoonotic disease spillover events.

@@

We extracted data from eligible studies into a Google sheets document. Extracted variables included i) study identifiers; ii) study aims; iii) trapping methodology; iv) geolocation data; v) method of speciation; vi) trapping locations and dates; vii) trapped species; viii) number of trap-nights and ix) microorganisms/pathogens of interest. The data extraction tool is archived and available in Supplementary 1 or at this [link](https://docs.google.com/spreadsheets/d/1rQYjHhk6uk1PoKZZVsgFlmuqWGicU2tTisk9ddfAwTM/edit?usp=sharing).

@@

We summarised the number of studies, the year in which trapping occurred and the country in which they were conducted. We used the GPS coordinates of single trap, trapping grid/line or study site and the number of trap nights to calculate trapping effort (trap night density) within level 2 administrative areas in West Africa. The median trap success rate for all rodents at a defined trap site was calculated separately for trap sites which included built environments and non-built environments due to important differences in trap success between these different environments. The number of rodents trapped at a trap site was then used in combination with these values of trap success to impute the number of trap nights for trap sites where there was no record of the number of trap nights. We summarised the habitat types of trap sites based on information reported in the study, the dictionary used to match habitat types is available in Supplementary 2.

For studies investigating rodent zoonoses we compared the location of trapping sites with SEDAC Global Population Density estimates for 2005 - the median year studies were started [@socioeconomic_data_and_applications_center_gridded_2021]. We used a Generalised Additive Model (GAM) incorporating a spatial interaction term to investigate the association of number of trap nights and human population density [@pedersen_hierarchical_2019]. The model structure was specified as:

$$
\text{Trap night density} \sim Tweedie(log(Population density_{2005}) + (Longitude * Latitude))   
$$

We obtained land cover classifications from the European Space Agency Copernicus dataset at 300m^2^ spatial resolution generated using Copernicus Climate Change Service information (2005) we extracted the proportion of land cover classes within all regions of West Africa and all regions in which rodent trapping occurred for investigation of zoonotic diseases. We compared the proportion of possible land cover classes with those of regions where trapping occurred.

Finally, we mapped the presence and absence of rodent species and compared this to the presence and absence produced from GBIF [@gbif_the_global_biodiversity_information_facility_occurrence_2021] and IUCN [@iucn_iucn_2021] datasets to give a measure of the extent of each species range in which they have been sampled.

@@

We identified 4,282 relevant citations, with 126 rodent trapping studies included. The list of included studies in available in Supplementary 3 or through the interactive web based application produced to explore this dataset, available [here](https://diddrog11.shinyapps.io/scoping_review_app/). The earliest included studies were published in 1974 with increasing numbers of studies being performed annually since 2005. The median length of rodent trapping activity was one year (IQR 0-2 years). The median time from completion of rodent trapping to publication is 3 years (IQR 2-5 years) (Figure 1.).

![Each row represents one of the 126 included studies, green points designate the first year of data collection, blue points designate the end of data collection. For studies completed within one year the blue point completely overlies the green. Studies with a transparent grey point did not report the year in which trapping was conducted. The year of publication is shown by a red point.](../figures/Fig_1_Panel_A.png)

@@

![**A:** Map of West Africa, countries where rodent trapping has occurred are mapped to level 2 administrative areas. The colour of the region corresponds to the trap night density per 1000 km^2. Colourless regions reported no trapping. **B:** This shows the linear predictor from obtained from the GAM. Predictions are limited to coordinates are trap sites (coloured areas), uncertainty in the linear predictor is represented by th transparency of the colour. Yellow regions represent areas with higher than expected trapping effort for the regional population and purple regions represent areas trapped at lower than expected rates based on their populations](../figures/Figure_2.png)

@@

![The proportion of the 10 land cover classes reported by the ESA Copernicus dataset at 300m^2 resolution are shown for all regions of West Africa (yellow) and for regions in which trapping occurred (purple). Rodent trapping occurred in regions over-representative for cropland, mosaic landscapes and shrubland wile occurring in regions under-representative for bare- and sparse vegetation land cover classes.](../figures/Figure_3.png)

@@

![Each row corresponds to a single rodent species. The column on the left shows the presence and absence of a rodent species from the individual studies included in this review. The centre column shows the presence of a rodent species obtained from GBIF (September 2021) for records where longitude and latitude have been provided. The right column shows the range of rodent species as proposed by the IUCN (2021) (red shaded area), overlaid are the presence points from both this review and GBIF records.](../figures/Figure_4.png)

@@

```{r, figure_5, echo=FALSE, out.width="100%", fig.cap="Presence/absence plots at each unique trapping site for the four most commonly assayed microorganisms Arenaviridae (top), *Bartonella sp.* (second row), *Borrelia sp.* (third row) and *Toxoplasma gondii* (fourth row). The tables to the right of each map highlight the 10 (or number applicable) most commonly positive and tested rodent species and genera assayed."}

knitr::include_graphics(here("figures", "Figure_5a.png"),
                        here("figures", "Figure_5b.png"),
                        here("figures", "Figure_5_legend.png"))

@@

@@

![For the 27 most commonly assayed rodent species and genera, the percentage of positive assays against each of the 32 microorganisms tested for is shown. The size of a point is proportional to the the number of individuals assayed for that microorganism.](../figures/Figure_6.png)

# Discussion

We identified 124 rodent trapping studies conducted in the West African region since 1965. These studies were conducted in 14 countries with 54 studies providing information about zoonotic pathogens in trapped rodents. Studies were conducted in all West African countries except Gambia and Togo with between- and within-country variability in the number of trap nights conducted. Over 70,000 rodents identified to genus or species level were trapped from 112 species of Rodentia. Thirty-two potential pathogens were investigated across all studies, 5 rodent species were assayed for at least 18 potential pathogens. We identify several important limitations to available data, potentially biasing previous zoonotic risk studies, first, in the timing of studies and the locations in which they've been conducted, second, the locations where a species was recored as present and third, the limited selection of pathogens assayed at these trapping sites.

First, the time period and location of trapping studies, when not systematically conducted, will introduce bias that must be accounted for when attempting to infer the risk of zoonotic disease spillover events or novel pathogen emergence. Most studies were conducted in a relatively short period since 2000. During this period human populations, agricultural farming and livestock intensification and urbanisation have progressed dramatically across West Africa [@acheampong_deforestation_2019; @connolly_extended_2021]. These changes alter the availability of habitats for rodent species, potentially increasing the prevalence of commensal species which are more likely to be reservoirs of zoonoses. Thus, incorporating the time period in which data were collected is important when modelling risk of zoonotic spillover or novel pathogen emergence based on known reservoir species distributions. In addition, several regions, namely North West Senegal, Eastern Sierra Leone, Southern Ghana, Southern Benin, South West and Central Nigeria have been sampled at much higher intensity than elsewhere in West Africa. This sampling heterogeneity will again introduce bias when inferring current or projected risk of zoonotic spillover or novel pathogen emergence.

Second, rodent species were found to have greater species ranges than those reported by IUCN data. Trapping studies were conducted from more regions across West Africa than were available from GBIF records. This increased spatial resolution and the inclusion of absence data alongside presence data presents a much more complete representation of the spatial heterogeneity of rodent species' locations within their ranges. Failing to incorporate this heterogeneity in favour of using ranges or presence only data to produce rodent reservoir species' distributions can importantly bias estimations of risk of zoonotic spillover or novel pathogen emergence. For example, the apparent absence of *Mastomys natalensis* at several trapping sites in Nigeria may explain some of the observed spatial variation in human Lassa fever cases [@redding_geographical_2021] that would not be expected based on this species IUCN range or presence only data from GBIF.

Third, while 32 potential zoonoses were investigated in at least one study, only four pathogen species or groups were investigated in more than three. Of these four pathogens, studies for *Lassa mammarenavirus* or other Arenaviridae where the most widely performed, both in absolute numbers but also by number of countries and number of study sites within countries. In contrast, *Toxoplasma sp.* were found to be prevalent in rodent hosts within Senegal - and despite a significant, known, burden of human disease - no studies investigated the prevalence of this pathogen in other West African countries [@bigna_global_2020]. The limited geographic scope of sampling for these pathogens hampers efforts to understand the risk and human health burden of zoonotic spillover across this diverse region.

Finally, we found that only five of the 112 rodent species trapped had been tested for more than half of the potential pathogens. There is increasing emphasis on the role of co-infection and multi-species pathogen transmission systems on zoonotic disease spillover risk which depends on comprehensive testing of both known and un-known reservoir species [@sweeny_synzootics_2021]. More comprehensive testing of potential reservoir species across West Africa is required to better understand the role of individual species in disease systems [@douglass_how_2016]. This will also reduce the risk of bias in predictive risk modelling by better characterising host-pathogen associations in species rich environments.

@@

Here we have identified important biases in the locations of studies, the trapping effort for different rodent species and the selection of zoonotic pathogens to assay across the West African region. The use of these data to infer risk of zoonotic spillover or novel disease emergence is likely inappropriate unless these systematic biases are addressed. We hope that this current synthesis of data will encourage further comprehensive ecological studies on rodent species assemblages and their pathogens. One priority would be to better understand the true distribution of known rodent reservoirs (particularly *Mastomys natalensis*) across the region with particular emphasis made to the habitat in which they reside and the structure of rodent assemblages.

@@

# Acknowledgments
The European Commission nor ECMWF are not responsible for any use that may be made of the Copernicus information or data it contains.

# Supplementary 1

**Data extraction tool**

# Supplementary 2

**Habitat dictionary**

# Supplementary 3

**Included studies table**
